Correct and Secure Web Programming using Dependent Types and Embedded Domain-Specific Languages

### 2013-07-25 paper/main.tex
Form construction / handling first draft completed

\titlebanner{Draft for submission to IFL 2013}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

\title{Correct and Secure Web Programming using 
Embedded Domain-Specific Languages}
%\subtitle{Using Embedded Domain-Specific Languages to Improve Confidence in Web Applications}

@@

An effect is represented as an instance of the \texttt{EFFECT} type. Such instances may be associated with a \textit{resource}, 
Describing a program in the \texttt{Effects} library is achieved using the following data type:
\begin{verbatim}
EffM : (m : Type -> Type) ->
       (es : List EFFECT) ->
       (es’ : List EFFECT) ->
       (a : Type) -> Type
\end{verbatim}

\texttt{EffM} is a type that is parameterised over an execution context (\texttt{m}), a list of effects that the program may use (\texttt{xs}), a list of effects at the end of the program's execution (\texttt{es'}), and a return type (\texttt{a}). The list of input types can be used to ensure that an operation may only be called when the program is in a particular state, and since the list of input effects may differ from the list of output effects, we may modify the resource associated with the state after each operation. This allows us to simply enforce a resource usage protocol, and deviation from this protocol will result in a compile-time error.

Implementing a new effect involves creating a new algebraic data type, of kind \texttt{Effect}, which describes the available operations and how they modify the resource associated with the effect. Each operation consists of an initial resource \texttt{res}, the resource at the end of the operation \texttt{res'}, and the output type of the operation \texttt{t}.

%TODO: Execution
\begin{verbatim}
Effect : Type
Effect = (res : Type) -> (res’ : Type) ->
         (t : Type) -> Type
\end{verbatim}

Operations are made available to users by writing functions with the correct \texttt{EffM} type, which may then be used to invoke the operation from within an effectful program.

\begin{verbatim}
queryPostVar : 
 String -> 
 Eff m [CGI (InitialisedCGI TaskRunning)] 
   (Maybe String)
queryPostVar query = (QueryPostVar query)
\end{verbatim}

Additionally, we define \textit{handlers}, which describe how each operation is interpreted in a given execution context. Finally, \texttt{Effect} instances are promoted to \texttt{EFFECT} instances through the use of the \texttt{MkEff} constructor, which is provided with the initial resource value.

@@

By modelling CGI as an algebraic effect, we may enforce a resource usage protocol which, even though the program may be running in an IO execution context, prevents arbitrary IO from being performed and therefore ensures that the headers are firstly written correctly. In order to accomplish this, we define an effect, \texttt{Cgi}, and an associated resource, \texttt{InitialisedCGI}, which is parameterised over the current state, \texttt{CGIStep}. This resource describes the current state, alongside a \texttt{CGIInfo} record which contains information from the request. We represent an uninitialised CGI effect as the unit type, ().

\begin{verbatim}
data CGIStep = Initialised 
             | TaskRunning 
             | TaskCompleted 
             | HeadersWritten 
             | ContentWritten

data InitialisedCGI : CGIStep -> Type where
  ICgi : CGIInfo -> InitialisedCGI s
\end{verbatim}

Figure ~\ref{fig:cgistates} shows the states through which the CGI program progresses. Each operation performed in an effectful program requires the resource to be of a certain type, and the completion of the operation may alter the resource.

@@

In order to make use of the CGI framework, users specify a function of the following type:
\begin{verbatim}
CGIProg : List EFFECT -> Type -> Type
CGIProg effs a = 
 Eff IO (CGI (InitialisedCGI TaskRunning) :: effs) a
\end{verbatim}
This is then passed, along with initial values for other effects that the user may wish to use, to the runAction function, which invokes the RunAction operation and executes the user-specified action.

@@

Figure ~\ref{fig:sqlitestates} shows the resource usage protocol for the SQLite library. There are three main phases involved in the usage of the SQLite effect: connection to the database, preparation of a query, and execution of the query. This is reflected in the resource associated with the SQLite effect, which again is parameterised by the current state.

\begin{verbatim}
data SQLiteRes : Step -> Type where
  OpenConn : DBPointer -> SQLiteRes s
  OpenStmt : DBPointer -> StmtPointer -> SQLiteRes s
  ExecutingStmt : DBPointer -> 
                  StmtPointer -> 
                  StepResult -> 
                  SQLiteRes s

@@

data StmtPointer = ValidStmt Ptr
                 | InvalidStmt 
\end{verbatim}
If a failure happens at any point during the computation, the resource is updated to reflect the failure. At this point, no further side-effecting requests are made to the underlying SQLite library, in order to ensure safety. The \texttt{connFail}, \texttt{stmtFail}, \texttt{bindFail} and \texttt{executeFail} utility functions allow for failures, once detected, to be handled by executing the appropriate sequence of state transition functions to dispose of any open resources and return to the initial state.

%Queries are evaluated through one or more calls to the \texttt{nextRow} function, which either executes an update statement or returns the next row of a result set. 
SQL queries are evaluated in SQLite upon a call to the library function \texttt{sqlite3\_step()}. In the case that a statement returns a result set, each subsequent call retrieves another row for processing using a column access function. Once all rows have been retrieved, the library returns \texttt{SQLITE\_DONE}, meaning that no further calls should be made without resetting the function. We encapsulate this requirement through the \texttt{StepResult} data type within the \texttt{ExecutingStmt} constructor.

\begin{verbatim}
data StepResult = Unstarted
                | StepFail
                | StepComplete
                | NoMoreRows
\end{verbatim}
Each call to \texttt{nextRow}, which is a wrapper around the underlying \texttt{sqlite3\_step()} function, returns a result of type \texttt{StepResult}, which is then reflected in the resource. Calls to \texttt{sqlite3\_step()} are only executed if the previous \texttt{StepResult} is either \texttt{Unstarted} or \texttt{StepComplete}. We may therefore statically guarantee that only calls that will return a valid result are executed.

@@

Firstly, we define a function of type:
\begin{verbatim}
Eff IO [SQLITE ()] (Either String (List (String, Int)))
\end{verbatim}
This means that the program will be run in the IO execution context, and must start and end with no active resources. The return type indicates that either a list of \text{(String, Int)} pairs, representing names and ages in the database, or an error will be returned.

\begin{verbatim}
testSelect : Eff IO [SQLITE ()] 
             (Either String (List (String, Int)))
testSelect = do
  open_db <- openDB "people.db"
  if open_db then do
    let sql = "SELECT * FROM `test`;"
    sql_prep_res <- prepareStatement sql
    if sql_prep_res then do 
      startBind
      finishBind
      beginExecution
      results <- collectResults
      finaliseStatement
      closeDB
      return $ Right results
    else do err <- stmtFail
            return $ Left err
  else do err <- connFail
          return $ Left err $
\end{verbatim}

The program firstly attempts to open a connection to the ``people.db'' database. At this point, since the \texttt{OpenDB} operation has been invoked, the program transitions to the \texttt{ConnectionOpened} state. The \texttt{openDB} function returns a Boolean value indicating whether or not the operation is successful. If not, then the \texttt{connFail} function is called to generate an appropriate error and dispose of the resources.

@@

Much akin to the SQLite effect, we encapsulate failure by reflecting it in the resource associated with the effect.

\begin{verbatim}
data SessionStep = SessionUninitialised
                 | SessionInitialised

data SessionRes : SessionStep -> Type where
  InvalidSession : SessionRes s  
  ValidSession   : SessionID -> 
                   SessionData -> 
                   SessionRes s
\end{verbatim}
The \texttt{SessionRes} data type is parameterised over the current state, which determines which operations may be performed, and has two constructors: \texttt{InvalidSession} and \texttt{ValidSession}. If an operation such as creating a new session fails, no further side-effecting calls will be made, in turn preserving integrity.

@@

\begin{verbatim}
SELECT `name`, `age` FROM `staff` LIMIT ' + page + ', 5
\end{verbatim}
The \texttt{page} variable is assumed to be an integer, but may instead be modified by an attacker to include a malicious string which would alter the semantics of the query, allowing an attacker to execute a blind SQL injection attack. % Might be a good idea to cite an SQL injection paper which uses LIMIT clauses here

@@

\subsection{Form Construction}
Each form element is specified to hold a particular type of data, which is then passed directly as that particular type, assuming that the correct type of data is specified by the user, to the handler function. In order to encapsulate this, we firstly define the allowed data types as part of an algebraic data type, \texttt{FormTy}.
\begin{verbatim}
data FormTy = FormString
            | FormInt
            | FormBool
            | FormFloat
            | FormList FormTy 
\end{verbatim}
In \idris{}, types are first-class, meaning that they may be treated like other terms in computations. We use this to convert between abstract and concrete representations of allowed form types:
\begin{verbatim}
interpFormTy : FormTy -> Type
interpFormTy FormString = String
interpFormTy FormInt = Int
interpFormTy FormBool = Bool
interpFormTy FormFloat = Float
interpFormTy (FormList a) = List (interpFormTy a)
\end{verbatim}
%
In order to specify a form, we once again make use of the \texttt{Effects} library in order to create a resource-aware EDSL. By recording the type of each form element as it is added in the type of the form, we may statically ensure that the user-supplied handler function is of the correct type to handle the data supplied by the form. If a handler function with with different types is supplied, then this will result in a compile-time type error.

@@

To specify a form instance, we define a function of type \texttt{UserForm}:
\begin{verbatim}
UserForm : Type
UserForm = EffM m [FORM (FormRes []) 
                        (FormRes [])] String
\end{verbatim}
All forms are required to include a submit button, as mandated by the requirement that the input and output resource contains an empty list of types; this requirement is fulfilled as per the output resource type of the \texttt{AddSubmit} operation. As the creation of a form is a pure function which does not include side effects, we do not restrict the handler to IO as with previously-discussed EDSLs, instead denoting the fact that it may be run in any handler with the implicit variable \texttt{m}.

@@

Before associating a handler function with the form, we must firstly specify the effects available to the handler. This is done through the use of the \texttt{useEffects}, which updates the list of effects in the type of the form resource. By doing this, we may subsequently use the effects in calculations at the type level, in particular when calculating the type of the handler function for the form.

\begin{verbatim}
useEffects : (effs : List WebEffect) ->
             EffM m [FORM (FormRes G E)] 
                    [FORM (FormRes G effs)] ()
useEffects effs = (UseEffects effs)
\end{verbatim}
At this point, it is unfortunately not possible to specify arbitrary effects due to difficulties with serialising initial resource environments. This is, however, not an inherent limitation as the \texttt{Effects} library allows for the introduction of effects within an effectful computation.
%
We may specify a handler function of type \texttt{FormHandler}:
\begin{verbatim}
FormHandler : List EFFECT -> Type
FormHandler effs = Eff IO effs ()
\end{verbatim}
In order to associate a handler with a form, we may call the \texttt{addSubmit} function:
%
\begin{verbatim}
  addSubmit : (f :  mkHandlerFn ((reverse G), E)) ->
              (fns : HandlerList) ->
              {default tactics { applyTactic 
                findFn 100; solve; }
               prf : FnElem f fns} ->
              EffM m [FORM (FormRes G E)]
                     [FORM (FormRes [] [])] 
                     String
  addSubmit f handlers {prf} = (Submit f name)
    where name : String
          name = getString' f handlers prf          
\end{verbatim}
%
Let us look at each aspect of this function in turn. Firstly, the \texttt{mkHandlerFn} function calculates the required type of the handler function from the list of types associated with the form elements, and the effects we specified with the \texttt{useEffects} function. Note that since we prepend types to the list of \texttt{FormTy}s as opposed to prepending them, we must reverse the list.
%
\begin{verbatim}
MkHandlerFnTy : Type
MkHandlerFnTy = (List FormTy, List WebEffect)

@@

mkHandlerFn : MkHandlerFnTy -> Type 
mkHandlerFn (tys, effs) = mkHandlerFn' tys effs 
\end{verbatim}
The \texttt{mkHandlerFn} function takes as its argument a tuple describing the arguments and web effects available to the function. When constructing the function type, we encase all arguments withing \texttt{Maybe} type, in order to handle failure should the supplied data fail to parse to the specified type. We may then extract the information through pattern matching.
%The type of a valid handler function is calculated by the \texttt{interpCheckedFnTy} function, given the vector of types expected by form elements, the return type of the handler function, and an abstract representation of effects permitted to be used by a handler function. 
%
%{\small
%\begin{verbatim}
%interpCheckedFnTy : Vect FormTy n -> 
%                    List EFFECT -> 
%                    Type -> 
%                    Type
%interpCheckedFnTy tys effs t = interpCheckedFnTy' (reverse tys)
% where interpCheckedFnTy' : Vect FormTy n -> Type
%       interpCheckedFnTy' [] = FormHandler effs t
%       interpCheckedFnTy' (x :: xs) = Maybe (interpFormTy x) ->
%                                      interpCheckedFnTy' xs
%\end{verbatim}
%}

@@

\subsection{Data Handling}
Once the form has been submitted, a web application may handle the submitted data by invoking the \texttt{HandleForm} CGI operation. This will firstly check for the existence of the hidden \texttt{handler} field, which contains the previously serialised metadata about the form handling function, before deserialising the data into a three-tuple of the argument types, required effects, and return type.

\subsection{Example Code}

\section{Extended Example: Message Board}
\section{Related Work}
\section{Conclusions}
\subsection{Further Work}
% Cryptography bindings would be really good.
% Integration with a web server instead of doing everything over CGI
% Improvements to form handling system

% JS backend

@@

\acks
% SICSA / 
%Acknowledgments, if needed.

>>>>

\titlebanner{Draft for submission to IFL 2013}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.
\title{Correct and Secure Web Programming using Dependent Types and
Embedded Domain-Specific Languages}
%\subtitle{Using Embedded Domain-Specific Languages to Improve Confidence in Web Applications}

@@

An effect is represented as an instance of the \texttt{EFFECT} type. Such instances may be associated with a \textit{resource}, 
Describing a program in the \texttt{Effects} library is achieved using the following data type:
{\small
\begin{verbatim}
EffM : (m : Type -> Type) ->
       (es : List EFFECT) ->
       (es’ : List EFFECT) ->
       (a : Type) -> Type
\end{verbatim}
}
\texttt{EffM} is a type that is parameterised over an execution context (\texttt{m}), a list of effects that the program may use (\texttt{xs}), a list of effects at the end of the program's execution (\texttt{es'}), and a return type (\texttt{a}). The list of input types can be used to ensure that an operation may only be called when the program is in a particular state, and since the list of input effects may differ from the list of output effects, we may modify the resource associated with the state after each operation. This allows us to simply enforce a resource usage protocol, and deviation from this protocol will result in a compile-time error.

Implementing a new effect involves creating a new algebraic data type, of kind \texttt{Effect}, which describes the available operations and how they modify the resource associated with the effect. Each operation consists of an initial resource \texttt{res}, the resource at the end of the operation \texttt{res'}, and the output type of the operation \texttt{t}.
{\small
%TODO: Execution
\begin{verbatim}
Effect : Type
Effect = (res : Type) -> (res’ : Type) ->
         (t : Type) -> Type
\end{verbatim}
}
Operations are made available to users by writing functions with the correct \texttt{EffM} type, which may then be used to invoke the operation from within an effectful program.
{\small
\begin{verbatim}
queryPostVar : 
 String -> 
 Eff m [CGI (InitialisedCGI TaskRunning)] 
   (Maybe String)
queryPostVar query = (QueryPostVar query)
\end{verbatim}
}
Additionally, we define \textit{handlers}, which describe how each operation is interpreted in a given execution context. Finally, \texttt{Effect} instances are promoted to \texttt{EFFECT} instances through the use of the \texttt{MkEff} constructor, which is provided with the initial resource value.

@@

By modelling CGI as an algebraic effect, we may enforce a resource usage protocol which, even though the program may be running in an IO execution context, prevents arbitrary IO from being performed and therefore ensures that the headers are firstly written correctly. In order to accomplish this, we define an effect, \texttt{Cgi}, and an associated resource, \texttt{InitialisedCGI}, which is parameterised over the current state, \texttt{CGIStep}. This resource describes the current state, alongside a \texttt{CGIInfo} record which contains information from the request. We represent an uninitialised CGI effect as the unit type, ().
{\small
\begin{verbatim}
data CGIStep = Initialised 
             | TaskRunning 
             | TaskCompleted 
             | HeadersWritten 
             | ContentWritten

data InitialisedCGI : CGIStep -> Type where
  ICgi : CGIInfo -> InitialisedCGI s
\end{verbatim}
}
Figure ~\ref{fig:cgistates} shows the states through which the CGI program progresses. Each operation performed in an effectful program requires the resource to be of a certain type, and the completion of the operation may alter the resource.

@@

In order to make use of the CGI framework, users specify a function of the following type:
{\small
\begin{verbatim}
CGIProg : List EFFECT -> Type -> Type
CGIProg effs a = 
 Eff IO (CGI (InitialisedCGI TaskRunning) :: effs) a
\end{verbatim}
}
This is then passed, along with initial values for other effects that the user may wish to use, to the runAction function, which invokes the RunAction operation and executes the user-specified action.

@@

Figure ~\ref{fig:sqlitestates} shows the resource usage protocol for the SQLite library. There are three main phases involved in the usage of the SQLite effect: connection to the database, preparation of a query, and execution of the query. This is reflected in the resource associated with the SQLite effect, which again is parameterised by the current state.
{\small
\begin{verbatim}
data SQLiteRes : Step -> Type where
  OpenConn : DBPointer -> SQLiteRes s
  OpenStmt : DBPointer -> StmtPointer -> SQLiteRes s
  ExecutingStmt : DBPointer -> 
                  StmtPointer -> 
                  StepResult -> 
                  SQLiteRes s

@@

data StmtPointer = ValidStmt Ptr
                 | InvalidStmt 
\end{verbatim}
}
If a failure happens at any point during the computation, the resource is updated to reflect the failure. At this point, no further side-effecting requests are made to the underlying SQLite library, in order to ensure safety. The \texttt{connFail}, \texttt{stmtFail}, \texttt{bindFail} and \texttt{executeFail} utility functions allow for failures, once detected, to be handled by executing the appropriate sequence of state transition functions to dispose of any open resources and return to the initial state.

%Queries are evaluated through one or more calls to the \texttt{nextRow} function, which either executes an update statement or returns the next row of a result set. 
SQL queries are evaluated in SQLite upon a call to the library function \texttt{sqlite3\_step()}. In the case that a statement returns a result set, each subsequent call retrieves another row for processing using a column access function. Once all rows have been retrieved, the library returns \texttt{SQLITE\_DONE}, meaning that no further calls should be made without resetting the function. We encapsulate this requirement through the \texttt{StepResult} data type within the \texttt{ExecutingStmt} constructor. 
{\small
\begin{verbatim}
data StepResult = Unstarted
                | StepFail
                | StepComplete
                | NoMoreRows
\end{verbatim}
}
Each call to \texttt{nextRow}, which is a wrapper around the underlying \texttt{sqlite3\_step()} function, returns a result of type \texttt{StepResult}, which is then reflected in the resource. Calls to \texttt{sqlite3\_step()} are only executed if the previous \texttt{StepResult} is either \texttt{Unstarted} or \texttt{StepComplete}. We may therefore statically guarantee that only calls that will return a valid result are executed.

@@

Firstly, we define a function of type:
{\small
\begin{verbatim}
Eff IO [SQLITE ()] (Either String (List (String, Int)))
\end{verbatim}
}
This means that the program will be run in the IO execution context, and must start and end with no active resources. The return type indicates that either a list of \text{(String, Int)} pairs, representing names and ages in the database, or an error will be returned. 
{\small
\begin{verbatim}
testSelect : Eff IO [SQLITE ()] 
             (Either String (List (String, Int)))
testSelect = do
  open_db <- openDB "people.db"
  if open_db then do
    let sql = "SELECT * FROM `test`;"
    sql_prep_res <- prepareStatement sql
    if sql_prep_res then do 
      startBind
      finishBind
      beginExecution
      results <- collectResults
      finaliseStatement
      closeDB
      return $ Right results
    else do err <- stmtFail
            return $ Left err
  else do err <- connFail
          return $ Left err $
\end{verbatim}
}
The program firstly attempts to open a connection to the ``people.db'' database. At this point, since the \texttt{OpenDB} operation has been invoked, the program transitions to the \texttt{ConnectionOpened} state. The \texttt{openDB} function returns a Boolean value indicating whether or not the operation is successful. If not, then the \texttt{connFail} function is called to generate an appropriate error and dispose of the resources.

@@

Much akin to the SQLite effect, we encapsulate failure by reflecting it in the resource associated with the effect. 
{\small
\begin{verbatim}
data SessionStep = SessionUninitialised
                 | SessionInitialised

data SessionRes : SessionStep -> Type where
  InvalidSession : SessionRes s  
  ValidSession   : SessionID -> 
                   SessionData -> 
                   SessionRes s
\end{verbatim}
}
The \texttt{SessionRes} data type is parameterised over the current state, which determines which operations may be performed, and has two constructors: \texttt{InvalidSession} and \texttt{ValidSession}. If an operation such as creating a new session fails, no further side-effecting calls will be made, in turn preserving integrity.

@@

{\small
\begin{verbatim}
SELECT `name`, `age` FROM `staff` LIMIT ' + page + ', 5
\end{verbatim}
}
The \texttt{page} variable is assumed to be an integer, but may instead be modified by an attacker to include a malicious string which would alter the semantics of the query, allowing an attacker to execute a blind SQL injection attack. % Might be a good idea to cite an SQL injection paper which uses LIMIT clauses here

@@

\subsection{Form Construction}
Each form element is specified to hold a particular type of data, which is then passed directly as that particular type, assuming that the correct type of data is specified by the user, to the handler function. In order to encapsulate this, we firstly define the allowed data types as part of an algebraic data type, \texttt{FormTy}.
{\small
\begin{verbatim}
data FormTy = FormString
            | FormInt
            | FormBool
            | FormFloat
            | FormList FormTy 
\end{verbatim}
}
In \idris{}, types are first-class, meaning that they may be treated like other terms in computations. We use this to convert between abstract and concrete representations of allowed form types:
{\small
\begin{verbatim}
interpFormTy : FormTy -> Type
interpFormTy FormString = String
interpFormTy FormInt = Int
interpFormTy FormBool = Bool
interpFormTy FormFloat = Float
interpFormTy (FormList a) = List (interpFormTy a)
\end{verbatim}
}
%
In order to specify a form, we once again make use of the \texttt{Effects} library in order to create a resource-aware EDSL. By recording the type of each form element as it is added in the type of the form, we may statically ensure that the user-supplied handler function is of the correct type to handle the data supplied by the form. If a handler function with with different types is supplied, then this will result in a compile-time type error.

@@

To specify a form instance, we define a function of type \texttt{UserForm}:
{\small
\begin{verbatim}
UserForm : Type
UserForm = EffM m [FORM (FormRes []) 
                        (FormRes [])] String
\end{verbatim}}
All forms are required to include a submit button, as mandated by the requirement that the input and output resource contains an empty list of types; this requirement is fulfilled as per the output resource type of the \texttt{AddSubmit} operation. As the creation of a form is a pure function which does not include side effects, we do not restrict the handler to IO as with previously-discussed EDSLs, instead denoting the fact that it may be run in any handler with the implicit variable \texttt{m}.

@@

Before associating a handler function with the form, we must firstly specify the effects available to the handler. This is done through the use of the \texttt{useEffects}, which updates the list of effects in the type of the form resource. By doing this, we may subsequently use the effects in calculations at the type level, in particular when calculating the type of the handler function for the form. 
{\small
\begin{verbatim}
useEffects : (effs : List WebEffect) ->
             EffM m [FORM (FormRes G E)] 
                    [FORM (FormRes G effs)] ()
useEffects effs = (UseEffects effs)
\end{verbatim}}
Whilst it is not possible to serialise arbitrary effects due to the associated difficulties with serialising initial resource environments, we allow for three effects to be serialised: \texttt{CGI}, \texttt{SQLITE} and \texttt{SESSION}. This is, however, not an inherent limitation as the \texttt{Effects} library allows for the introduction of effects within an effectful computation.
%
We may specify a handler function of type \texttt{FormHandler}:
\begin{verbatim}
FormHandler : List EFFECT -> Type
FormHandler effs = Eff IO effs ()
\end{verbatim}
In order to associate a handler with a form, we may call the \texttt{addSubmit} function:
%
\begin{verbatim}
  addSubmit : (f :  mkHandlerFn ((reverse G), E)) ->
              (fns : HandlerList) ->
              {default tactics { applyTactic 
                findFn 100; solve; }
               prf : FnElem f fns} ->
              EffM m [FORM (FormRes G E)]
                     [FORM (FormRes [] [])] 
                     String
  addSubmit f handlers {prf} = (Submit f name)
    where name : String
          name = getString' f handlers prf          
\end{verbatim}
%
Let us look at each aspect of this function in turn. Firstly, the \texttt{mkHandlerFn} function calculates the required type of the handler function from the list of types associated with the form elements, and the effects we specified with the \texttt{useEffects} function. Note that since we prepend types to the list of \texttt{FormTy}s as opposed to prepending them, we must reverse the list.
{\small
\begin{verbatim}
MkHandlerFnTy : Type
MkHandlerFnTy = (List FormTy, List WebEffect)

@@

mkHandlerFn : MkHandlerFnTy -> Type 
mkHandlerFn (tys, effs) = mkHandlerFn' tys effs 
\end{verbatim}
}
The \texttt{mkHandlerFn} function takes as its argument a tuple describing the arguments and web effects available to the function. When constructing the function type, we encase all arguments withing \texttt{Maybe} type, in order to handle failure should the supplied data fail to parse to the specified type. We may then extract the information through pattern matching.

To store a reference to a handler function, we use the \texttt{HandlerFn} type:
\begin{verbatim}
public
HandlerFn : Type
HandlerFn = (ft ** (mkHandlerFn ft, String))
\end{verbatim}
%
The \texttt{**} notation denotes a dependent pair, wherein one argument, in this case the concrete handler function, depends on another, namely \texttt{MkHandlerFnTy} data used to construct the type of the handler function. We also store a unique string identifier, which is used to serialise a reference to the handler function. We then store each \texttt{HandlerFn} in a \texttt{HandlerList}.

\begin{verbatim}
public
HandlerList : Type
HandlerList = List HandlerFn
\end{verbatim}
To enforce the requirement that a supplied handler function must reside in the list of available handlers, and therefore allow us to retrieve the name with which to serialise the handler, we require a \textit{list membership proof},  \texttt{FnElem f fns}, which statically guarantees that a given item resides in a list.
{\small
\begin{verbatim}
  data FnElem : mkHandlerFn ((reverse G), E) -> 
                HandlerList -> Type where

       FnHere : {xs : HandlerList, f : 
                 mkHandlerFn ((reverse G), E)} ->
         FnElem f ((((reverse G), E) ** (f, fStr)) :: xs)
       FnThere : {xs : HandlerList, f : 
                 mkHandlerFn ((reverse G), E)} ->
               FnElem f xs -> FnElem f (x :: xs)
\end{verbatim}
}
\texttt{FnElem} is parameterised over \texttt{G} and \text{E}, the types of the form elements and the effects used by the handler function. \texttt{FnHere} demonstrates that the element is at the head of the current point of the list, whereas \texttt{FnThere} demonstrates that the element is at some point further in the list. %TODO: this should probably be rewritten
We may then use linguistic reflection and a simple automated proof search to automatically generate the proof at compile time, should one exist. The proof may then be used in subsequent computations: in our case, we use it to retrieve the unique identifier for the function. If the automated proof search fails, compilation will fail with a compile error.

@@

\subsection{Form Handling}
Once the form has been submitted, a web application may handle the submitted data by invoking the \texttt{HandleForm} CGI operation. This will firstly check for the existence of the hidden \texttt{handler} field, which contains the previously serialised metadata about the form handling function, before deserialising the data into a \texttt{MkHandlerFnTy}.

With this data, we may then look up the function in the list of registered handlers by using the serialised unique handler identifier. In order to apply the handler function to the data submitted in the form, we must firstly prove to the type-checker that the deserialised \texttt{MkHandlerFnTy} is the same as the one retrieved from the list of registered handlers. We do this by making use of the \texttt{decEq} function, which determines whether two types are equal, returning a proof of equality if so, and a proof of inequality if not.

With this proof, we may then use the \texttt{with} construct, inspired by \textit{views} in Agda \cite{viewfromleft}, to rewrite the arguments on the left hand side. This allows us to construct a function which, given the handler stored in the list of handlers, the data required to construct the function type and the \texttt{MkHandlerFnTy} deserialised from the form, determines whether the two \texttt{MkHandlerFnTy}s are decidably equal. If so, we may rewrite this on the left hand side, and therefore demonstrate that the recorded function may also be used to handle the form data. If not, the computation is abandoned and an error is shown. % I gave it a shot, right? :P
{\small
\begin{verbatim}
checkFunctions : (reg_fn_ty : MkHandlerFnTy) -> 
                 (frm_fn_ty : MkHandlerFnTy) -> 
                 mkHandlerFn reg_fn_ty -> 
                 Maybe (mkHandlerFn frm_fn_ty)
checkFunctions reg_ty frm_ty reg_fn with 
                             (decEq reg_ty frm_ty)
  checkFunctions frm_ty frm_ty reg_fn | Yes refl = Just reg_fn
  checkFunctions reg_ty frm_ty reg_fn | No _ = Nothing
\end{verbatim}}
We may then parse the arguments according to the types specified by the handler function, and then apply the arguments to the handler function through the use of partial application. Finally, we may run the handler function, ensuring that all updates made to the CGI state are propagated.
\section{Extended Example: Message Board}
\section{Related Work}
\section{Conclusions}
\subsection{Further Work}
% Cryptography bindings would be really good.
% Integration with a web server instead of doing everything over CGI
% Improvements to form handling system

@@

\acks
% SICSA / EPSRC (grant number?)
% #idris
% Idris contributors, in particular Melissa for the SQLite bindings and whoever wrote Network.Cgi
